Textmining in Economics

### 2017-10-17 textm.Rmd
little bit more on slides

```{r, include=FALSE}
rm(list = ls())
load(file = "~/GitHub/politsentiment_data/out/tweets_18_24.rda")
# during and after elecction
tweets <- tweets[which(tweets$day == "24"),]
tweets <- with(tweets, tweets[hour(date) >= 12 & hour(date) <= 23, ])
```

@@

1. Pre-process the text (Stopword removal, tokenization, stemming) and Represent(**document** $D$ as a numerical array $\boldsymbol{C}$ (e.g. *bag-of-words model*)

@@

## Cleaning Data

```{r, echo=FALSE}
htmlTable(orig[1:4,c("text","text_cleaned")], align = "l")
```

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
orig$text_cleaned <- removeWords(orig$text_cleaned, c("cdu","csu","union","spd","fdp","grüne","grünen","linke","afd"))

orig %>%
  #group_by(partei) %>%
  unnest_tokens(word, text_cleaned) %>%
  count(word, sort = TRUE)  %>%
  top_n(15) %>%
  htmlTable()
```

## Tokenization (Bigrams)

```{r, echo=FALSE, message=FALSE, warning=FALSE}
orig %>%
  #group_by(partei) %>%
  unnest_tokens(bigram, text_cleaned, token="ngrams", n=2) %>%
  count(bigram, sort = TRUE)  %>%
  top_n(15) %>%
  htmlTable()
```

@@

***
How to reflect the importance of a term? 3 concepts:

- TF: Term-Frenquency $tf_{v,d}$ for term $v$ and document $d$
- IDF: Inverse Document Frequency $idf_v = log(D/df_v)$, where $df_v$ is the number of documents that contain term $v$ and $D$ ist die number of Documents
- TF-IDF: $tf-idf_{d,v}=tf_{d,v} * idf_v$
  - Gives prominence to words that occur many times in few documents.

```{r, include=FALSE}
news.token <- tweets %>%
  group_by(partei) %>%
  unnest_tokens(word, text_cleaned) %>%
  count(partei, word, sort = TRUE)  %>%
  bind_tf_idf(word, partei, n) %>%
  arrange(desc(tf_idf))
```

```{r, echo=FALSE, comment = ""}
news.token[1:10,]
```

@@

  - The function relating $\boldsymbol{c}_i$ to $\boldsymbol{v}_i$ is unknown, as we cannot observe the true value of $v_i$. Unsupervised machine learning involves taking unclassified observations and uncovering hidden patterns that structure them in some meaningful way.

  - The outputs of algorithms for unsupervised machine learning can be used as inputs into econometric models for predicting some variable of interest, but this is a different approach from intentionally choosing the dimensions of content based on their predictive ability.

  - E.g.: Principal Component Analysis, Topic Modeling (latent Dirichlet Allocation)

@@

method for finding a group of words (i.e topic) from a collection of documents that best represents the information in the collection.

  - Discovering hidden topical patterns that are present across the collection
  - Annotating documents according to these topics
  - Using these annotations to organize, search and summarize texts

There are many techniques that are used to obtain topic models: Latent Dirichlet allocation

>>>>

```{r, include=FALSE}
rm(list = ls())
load(file = "~/GitHub/politsentiment_data/out/tweets_18_24.rda")
# during and after elecction
tweets <- tweets[which(tweets$day == "24"),]
tweets <- with(tweets, tweets[hour(date) >= 12 & hour(date) <= 23, ])

col <- c("deepskyblue1", "black","limegreen", "darkorchid2","gold", "red2")
names(col) <- levels(tweets$partei)
```

@@

1. **Pre-process** the text (Stopword removal, tokenization, stemming) and represent document $D$ as a numerical array $\boldsymbol{C}$ (e.g. bag-of-words model)

@@

## Cleaning Data (Example)

```{r, echo=FALSE}
orig$retweet_count <- as.numeric(orig$retweet_count)

orig %>%
  arrange(desc(retweet_count)) -> orig

htmlTable(orig[c(1,4,5,7),c("text","text_cleaned")], align = "l")
```

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
orig$text_cleaned <- removeWords(orig$text_cleaned, c("cdu","csu","union","spd","fdp","grüne","grünen","linke","afd","btw"))

orig %>%
  #group_by(partei) %>%
  unnest_tokens(word, text_cleaned) %>%
  count(word, sort = TRUE) -> tokens

orig %>%
  #group_by(partei) %>%
  unnest_tokens(bigram, text_cleaned, token="ngrams", n=2) %>%
  count(bigram, sort = TRUE) -> bigrams
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
tokens %>%
  top_n(20) %>%
  ggplot() +
  aes(x= reorder(word, n), y=n) +
  geom_col() +
  labs(title = "Tokens") +
  xlab("") +
  coord_flip() -> p1

bigrams %>%
  top_n(20) %>%
  ggplot() +
  aes(x= reorder(bigram, n), y=n) +
  geom_col() +
  labs(title = "Bigrams") +
  xlab("") +
  coord_flip() -> p2

grid.arrange(p1, p2, ncol = 2)
```

## Wordclouds
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(wordcloud)
# WordClouds ####
# get text
cdu <- orig[which(orig$partei=="CDU"),]
spd <- orig[which(orig$partei=="SPD"),]
fdp <- orig[which(orig$partei=="FDP"),]
gruene <- orig[which(orig$partei=="Die Grünen"),]
linke <- orig[which(orig$partei=="DIE LINKE"),]
afd <- orig[which(orig$partei=="AfD"),]

# Join texts in a vector 
cdu <- paste(cdu$text_cleaned, collapse = " ")
spd <- paste(spd$text_cleaned, collapse = " ")
fdp <- paste(fdp$text_cleaned, collapse = " ")
gruene <- paste(gruene$text_cleaned, collapse = " ")
linke <- paste(linke$text_cleaned, collapse = " ")
afd <- paste(afd$text_cleaned, collapse = " ")

# put everything in a single vector
all <- c(cdu, spd, fdp,gruene, linke, afd)

ignore <- paste(c("cdu","bündnis", "spd","fdp","gruene","grüne","grünen", "linke","dielinke","afd","gen","merkel","schulz","csu","union"), collapse = "|")
all <- gsub(ignore,"", all)

# Step 2: Corpus and term-docs_cleaned matrix ####
# create corpus
corp <- Corpus(VectorSource(all))

# create term-docs_cleaned matrix
tdm <- TermDocumentMatrix(corp)

# convert as matrix
tdm <- as.matrix(tdm)

# add column names
colnames(tdm) <- c("CDU","SPD","FDP","Die Grünen", "DIE LINKE","AfD")

# Step 3: Plot comparison wordcloud ####
comparison.cloud(tdm, random.order=FALSE, colors = c("black","red2", "gold","limegreen", "darkorchid2","deepskyblue1"), scale=c(3,.5),
                 title.size=1.2, max.words=100)
```

@@

## Term Frequency

How to reflect the importance of a term? 3 concepts:

- **TF** Term-Frenquency $tf_{v,d}$ for term $v$ and document $d$
- **IDF** Inverse Document Frequency $idf_v = log(\frac{D}{df_v})$, where $df_v$ is the number of documents that contain term $v$ and $D$ ist die number of Documents
- **TF-IDF** $tf-idf_{d,v}=tf_{d,v} * idf_v$
  - Gives prominence to words that occur many times in few documents.

## Term Frequency (Example)

```{r, include=FALSE}
tokens <- orig %>%
  group_by(partei) %>%
  unnest_tokens(word, text_cleaned) %>%
  count(partei, word, sort = TRUE)  %>%
  bind_tf_idf(word, partei, n)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
tokens %>%
  top_n(5) %>%
  ggplot() +
  aes(x= reorder(word, tf_idf), y=tf_idf, fill = partei) +
  geom_col(alpha=.8, show.legend = FALSE) +
  labs(x = NULL, y = "tf-idf") +
  scale_fill_manual(name="", values=col) +
  coord_flip() -> p1

tokens %>%
  top_n(5) %>%
  ggplot() +
  aes(x= reorder(word, tf), y=tf, fill = partei) +
  geom_col(alpha=.8, show.legend = FALSE) +
  labs(x = NULL, y = "tf") +
  scale_fill_manual(name="", values=col) +
  coord_flip() -> p2

grid.arrange(p1, p2, ncol = 2)
```

@@

  - **Starting Point** The true value of $v_i$ is unobserved -> The function relating $\boldsymbol{c}_i$ to $\boldsymbol{v}_i$ is unknown

  - **Idea** Uncover hidden patterns that structure the observations in some meaningful way.

  - **Use** The outputs of algorithms can be used as inputs into econometric models for predicting some variable of interest.

  - **E.g.** Principal Component Analysis, Topic Modeling (latent Dirichlet Allocation)

@@

A method for finding a group of words (i.e topic) from a collection of documents that best represents the information in the collection.

  - Discovering **hidden topical patterns** that are present across the collection
  - Annotating documents according to these topics
  - Using these annotations to organize, search and summarize texts

Very popular topic model is the **Latent Dirichlet allocation**
